---
title: SticsRFiles
output: 
  github_document:
   html_preview: true     
        
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
 
<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![R build status](https://github.com/SticsRPacks/SticsRFiles/workflows/R-CMD-check/badge.svg)](https://github.com/SticsRPacks/SticsRFiles/actions)
[![Codecov test coverage](https://app.codecov.io/gh/SticsRPacks/SticsRFiles/branch/master/graph/badge.svg)](https://app.codecov.io/gh/SticsRPacks/SticsRFiles?branch=master)

[![DOI](https://zenodo.org/badge/187986787.svg)](https://zenodo.org/badge/latestdoi/187986787)
<!-- badges: end -->

The goal of SticsRFiles is to perform manipulations of the [STICS](https://www6.paca.inrae.fr/stics_eng/)  model files either on XML files (used by the JavaStics GUI) or on text files used by the model fortran executable.

The basic functionnalites allows to read parameters names and values through XML queries and replace parameters values in files. A starting guide is available on the [Get started page](https://sticsrpacks.github.io/SticsRFiles/articles/SticsRFiles.html).

Advanced functionnalities are dedicated to produce XML parameter files using a mailing process like from XML templates and Excel sheets containing multiple simulations contexts. A JavaStics workspace is generated and directly usable from the JavaStics interface (GUI or command line), or with an R JavaStics interface provided by the [SticsOnR](https://SticsRPacks.github.io/SticsOnR) package.

If you want to be notified when a new release of this package is made, you can tick the Releases box in the "Watch / Unwatch => Custom" menu at the top right of [this page](https://github.com/SticsRPacks/SticsRFiles).


## Prerequisites and technical tips

> Some information about software requirements and operating system constraints are given in the SticsOnR package documentation [here](https://sticsrpacks.github.io/SticsOnR/).

## Installation

---

_**Warning:**_ *If during the installation process, packages updates are suggested.*

* First, abort the installation.
* Second, update the installed packages, except the `XML` package.
* Finally, install the SticsRFiles package.

---

### Recommended installation: `SticsRPacks`

The best way to install the packages from `SticsRPacks`, from which `SticsRFiles` is part of, is by installing the `[SticsRPacks]` package. The package can be installed from [GitHub](https://github.com/) using either the `devtools` package, or the more lightweight `remotes` package: 

```{r eval=FALSE}
devtools::install_github("SticsRPacks/SticsRPacks")
```

* With `remotes`

```{r eval=FALSE}
remotes::install_github("SticsRPacks/SticsRPacks")
```

The package will install the packages for you at the latest release version. 

### Or, install directly 

* With `devtools`

```{r eval=FALSE}
devtools::install_github("SticsRPacks/SticsRFiles@*release")
```

* With `remotes`

```{r eval=FALSE}
remotes::install_github("SticsRPacks/SticsRFiles@*release")
```

Normaly, all the package dependencies will be installed for CRAN packages.

## Examples

### Files manipulations
* A description of how to use the functions for manipulating XML input files is detailed 
[here](https://sticsrpacks.github.io/SticsRFiles/articles/Manipulating_Stics_XML_files.html)

* A description of how to use the functions for manipulating text input files is available [here](https://sticsrpacks.github.io/SticsRFiles/articles/Manipulating_Stics_text_files).

### Files generation
* A description of usefull functions for generating XML input files from usms parameters data stored in Excel files sheets or CSV files is available [here](https://sticsrpacks.github.io/SticsRFiles/articles/Generating_Stics_XML_files.html).

* A description of functions for generating text input files from XML one of usms parameters is available [here](https://sticsrpacks.github.io/SticsRFiles/articles/Generating_Stics_text_files.html).

## Getting help

If you have any question or suggestion or if you want to report a bug, please do it via the GitHub [issues](https://github.com/SticsRPacks/SticsRFiles/issues).

Thanks for that, this would greatly help us to improve this package.

## Citation

If you have used this package for a study that led to a publication or
report, please cite us. You can either use the citation tool from Github if you used the last version, or use `citation("SticsRFiles")` from R otherwise.

## Code of conduct

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.

The package is under intensive development, so you can fill an issue or request us a feature [here](https://github.com/SticsRPacks/SticsRFiles/issues) at any time. 

