---
title: "Generating Stics XML files from XL files data, `mailing` use case"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating_Stics_XML_files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup, echo=FALSE, eval = TRUE}
suppressMessages(library(SticsRFiles))
```


```{r, echo = FALSE, eval = TRUE}
workspace_path <- tempdir()
```

# Introduction

* In this part information are focusing on functionnalities related to Stics XML files generation as `mailing` capabilities.

* The generation of wheather files files are not included yet in the package, but will be included in a future version. One can use included toll in JavaStics GUI to perform thes files generation.

* The functions operate on data coming stored in data.frames loaded from Excel sheets



# Package functions


  * Generate xml files using a table of parameters     
    * gen_ini_xml
    * gen_sta_xml
    * gen_tec_xml
    * gen_usms_xml
    * gen_sols_xml

      
  * Used for manipulation of data from tables
    * get_params_dict


# Getting example files and scripts from the library

## Getting inputs XL example file: 

```{r}
# Getting the file in the current directory
copy_mailing_example()
# getting the file in a specic directory
copy_mailing_example(dest_dir = "/path/to/dest/dir")

#> [1] "inputs_stics_example.xlsx  has been copied in directory  
#> "/path/to/dest/dir

```

```{r, eval=TRUE, echo = FALSE, results='hide'}
workspace_path
copy_mailing_example(dest_dir = workspace_path)
```

## Getting example scripts for producing ini, sols, sta, tec and usms from the package
### Getting scripts files list     
```{r, eval=TRUE}
library(SticsRFiles)
copy_mailing_script_example()

```      
### Getting a copy of example scripts
  
  The files may be copied either to the current directory (i.e. returned by `getwd()`) or in a specific directory
  given as `dest_dir`.
  
```{r}
# a specific
copy_mailing_script_example("generate_ini_files.R")
copy_mailing_script_example("generate_ini_files.R", dest_dir = "/path/to/dest/dir")

# all scripts
copy_mailing_script_example("all")
copy_mailing_script_example("all", dest_dir = "/path/to/dest/dir")
```

```{r, eval=TRUE, echo = FALSE, results='hide'}
#workspace_path
copy_mailing_script_example("generate_usms_file.R", dest_dir = workspace_path)
```
### Example of a script content for generating the usms XML file (`generate_usms_file.R`)

```{r generate_usms_file}
# Generate an usms.xml file

library(readxl)
library(SticsRFiles)

out_file <- "/path/to/file/usms.xml" # or something like C:/path/to/file/usms.xml" for Windows
xl_dir <- "/path/to/xl/dir"          # or something like C:/path/to/xl/dir" for Windows

# Using an XL example file (from the package)
copy_mailing_example("inputs_stics_example.xlsx", dest_dir = xl_dir)
xl_path <- file.path(xl_dir, "inputs_stics_example.xlsx")

# Reading the xl file
xl_param <- read_excel(xl_path,sheet = "USMs")

# Generating a new usms.xml file, for all xl_param lines
gen_usms_xml(usms_param = xl_param, usms_out_file = out_file)
```


```{r eval_generate_usms_file, eval = TRUE, echo = FALSE, results='hide'}
# Generate an usms.xml file

library(readxl)
library(SticsRFiles)

out_file <- file.path(workspace_path,"usms.xml")
xl_dir <- workspace_path

# Using an XL example file (from the package)
copy_mailing_example("inputs_stics_example.xlsx", dest_dir = xl_dir)
xl_path <- file.path(xl_dir, "inputs_stics_example.xlsx")

# Reading the xl file
xl_param <- read_excel(xl_path,sheet = "USMs")

# Generating a new usms.xml file, for all xl_param lines
gen_usms_xml(usms_param = xl_param, usms_out_file = out_file)
```
### Adaptation of the script content

   * Only adapt the paths / directories and files names. In these scripts the XL file is given just as an example. It will be replaced by the user ones.
   
   * The table loaded from an XL sheet is entirely used in the examples, but a selection may be made to generate files only for a subset
   

# Constraints or limitations
  
## For `usms` and `sols` files: 
       
* For the moment the functions are not taking into account the xml files content update: i.e. parameters values modification for existing `usm` or `sol`, or adding new usm or sol. This means that xml sols or usms files are newly generated upon values contained in the xl sheets. 
* So, all the **usm** or **sol** parameters must be present in the XL sheet
    
## For `tec`, `sta` and `ini` files : 
    
* Take care with consistency of the base xml file provided to be duplicated and modified with parameters present in XL sheets. Messages will be displayed in case of a misspelled parameter, or an unknown parameter in the xml file which may correspond to another Stics version than the desired one.
  
* All the parameters are not mandatory in the XL sheet, only those that are to be modified



## Note on parameters names syntax
* The exact parameters names existing in xml files (with respect to case sensitive) must be used generally in XL sheets. But, we added some **shortcuts parameters** names for some of the which are too long or meaningless for the user.   

* A list have been added to facilitate xl sheet filling. It have been included while retreiving data from data frame with names replacement for existing ones in the correspondence list.     
At the moment, this list is very short and restricted to few `crop management parameters` (see table below). But we intent to generalize the use of that kind of correspondance list to facilitate comprehension and parameters names manipulations.
  
```{r echo=FALSE, warning=FALSE, message=FALSE, eval = TRUE}
library(knitr)
library(SticsRFiles)
dict <- get_params_dict()
dt <- data.frame(shortcuts=names(dict), real_name=unlist(dict,use.names = FALSE))

kable(dt, caption = "List of possible builtin parameters names substitution")
```

* To get this correspondance list : call get_params_dict(), as follows

```{r, eval = TRUE}
library(knitr)
library(SticsRFiles)
get_params_dict()

```

* But, it is possible to create and provide a personal list through the `dict` argument of the function.
This argument is usable at the moment for producing only *tec* files.  When calling `gen_tec_xml` the argument dict may be set as follows, for example for irrigations parameters:

```{r}  

irr_dict <- list(IrrDay="julapI_or_sum_upvt", IrrDose="amount") 
  
gen_tec_xml(..., dict = irr_dict) 
```

* The final used dictionnary is calculated as follows, merged from the internal one and the given one 
```{r, eval=TRUE}  

irr_dict <- list(IrrDay="julapI_or_sum_upvt", IrrDose="amount") 
  
dict <- get_params_dict(irr_dict)

dict

```

## Files names or usms/sols names in xl sheet
  * key words for detecting names or file names column: usm_, tec_, ini_, soil_, sta_

  * files names : may be either full files names or only file names without any suffix like _tec.xml, _ini.xml or _sta.xml 


## Consistency data checks in XL sheets    
For the moment *no consistency checks* functionnalities have been included in the package. So, we must take care of links between xml files names for example files names defined for usms related to names declared in `Tec`, `Station`, and so on. 


## XL sheets names    
XL sheet names are completely free, but must be meaningfull regarding to the kind of parameters they contain.

