---
title: "Generating Stics text files from XML files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating_Stics_text_files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params:
  javastics_path: /tmp/JavaSTICS-1.41-stics-9.1
  workspace_path: /tmp/JavaSTICS-1.41-stics-9.1/example
  output_path: /tmp/example
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup, echo = TRUE, eval = TRUE}
library(SticsRFiles)
library(SticsOnR)
```


```{r, echo = FALSE, eval = TRUE}
javastics_path <- params$javastics_path
workspace_path <- params$workspace_path
output_path <- params$output_path
```

## Introduction

* In this part information are focusing on functionalities related to generation of text input files for the Stics model.

* Generated files will be directly usable for running the model with **SticsOnR** package functions. See the documentation [here](https://sticsrpacks.github.io/SticsOnR).


## Converting JavaStics workspace files
For using the model directly neither using the JavaStics graphical interface nor the `run_javastics` function interface, we provide a function, `gen_usms_xml2txt`, for converting JavaStics XML files to Stics text files from a JavaStics workspace. 

Then, the model can be run using the [`run_stics()`](https://sticsrpacks.github.io/SticsOnR/reference/run_stics.html) function from [SticsOnR](https://sticsrpacks.github.io/SticsOnR/index.html)

Observation files may also be copied if they have a standard name corresponding to a usm name and a `.obs` extension. If not, they must be renamed to do so.

```{r paths_def, eval = FALSE}
# Specifying the JavaStics folder
javastics_path <- "/path/to/JavaSTICS-1.41-stics-9.1"

# Specifying a workspace as a subfolder of JavaStics 
workspace_path <- "example"
# or an absolute path to an external folder
# workspace_path <- "/path/to/javastics/workspace"

# Specifying an output folder path 
output_path <- "/path/to/output/folder"

```

## Converting files into separated folders (one per usm)
```{r convert_to_separate_1, eval = FALSE}
## Generating files for all the usms contained in the workspace

# Into the workspace directory
gen_usms_xml2txt(javastics_path, workspace_path)
#> ℹ USM 'SugarCane' successfully created
#> ℹ USM 'potato' successfully created
#> ℹ USM 'banana' successfully created
#> ℹ USM 'sorghum' successfully created
#> ℹ USM 'barley' successfully created
#> ℹ USM 'sugarbeet' successfully created
#> ℹ USM 'wheat' successfully created
#> ℹ USM 'maize' successfully created
#> ℹ USM 'soybean' successfully created
#> ℹ USM 'lettuce' successfully created
#> ℹ USM 'tomato' successfully created
#> ℹ USM 'DurumWheat' successfully created
#> ℹ USM 'rapeseed' successfully created
#> ℹ USM 'sunflower' successfully created
#> ℹ USM 'grass' successfully created
#> ℹ USM 'BareSoil' successfully created
#> ℹ USM 'demo_Wheat1' successfully created
#> ℹ USM 'demo_BareSoil2' successfully created
#> ℹ USM 'demo_maize3' successfully created
#> ℹ USM 'DurumWheat_4years' successfully created
#> ℹ USM 'maize_4years' successfully created
#> ℹ USM 'strawberry' successfully created
#> ℹ USM 'pea' successfully created
#> ℹ USM 'vine' successfully created
#> ℹ USM 'fescue' successfully created
#> ℹ USM 'flax' successfully created
#> ℹ USM 'intercrop_pea_barley' successfully created
#> ℹ USM 'timothy' successfully created
#> ℹ USM 'DurumWheat_snow' successfully created
#> ℹ USM 'Turmeric' successfully created
#> ℹ USM 'cc_BristleOat' successfully created
#> ℹ USM 'cc_mustard' successfully created
#> ℹ USM 'cc_ItalianRyegrass' successfully created
#> ℹ USM 'cc_vetch' successfully created
#> ℹ USM 'cc_CrimsonClover' successfully created
#> ℹ USM 'proto_rice' successfully created

```

```{r convert_to_separate_2, eval = FALSE}

# Into a specific output folder, with verbose mode turned off
gen_usms_xml2txt(javastics_path, workspace_path,
                 target_path = output_path,
                 verbose = FALSE)


## Generating files for a subset of usms
# Into the workspace directory
gen_usms_xml2txt(javastics_path, workspace_path, 
                 usms_list = c("banana","wheat"))
#> ℹ USM 'banana' successfully created
#> ℹ USM 'wheat' successfully created

# Into a specific folder
gen_usms_xml2txt(javastics_path, workspace_path, 
                 usms_list = c("banana","wheat"),
                 target_path = output_path)
#> ℹ USM 'banana' successfully created
#> ℹ USM 'wheat' successfully created

```

```{r convert_to_separate_3, eval = FALSE}
## Getting returned information about files generation
gen_info <- gen_usms_xml2txt(javastics_path, workspace_path,
                             usms_list = c("banana","wheat"),
                             target_path = output_path)
#> ℹ USM 'banana' successfully created
#> ℹ USM 'wheat' successfully created


gen_info
#> $usms_path
#> [1] "/path/to/JavaSTICS-1.41-stics-9.1/example/banana"
#> [2] "/path/to/JavaSTICS-1.41-stics-9.1/example/wheat" 
#> 
#> $files
#>  [1] "climat.txt"      "param.sol"       "ficini.txt"      "ficplt1.txt"    
#>  [5] "fictec1.txt"     "station.txt"     "new_travail.usm" "tempopar.sti"   
#>  [9] "tempoparv6.sti"  "ficplt2.txt"     "fictec2.txt"    
#> 
#> $copy_status
#> [1] TRUE TRUE
#> 
#> $obs_copy_status
#> [1] TRUE TRUE
```

## Converting files into one folder (overwriting case) 
```{r convert_to_one}
# Generating files directly in the workspace or a specific folder (no usm sub-folder) 
# In this case the model files are overwritten at each gen_usms_xml2txt call !
# In the workspace
gen_usms_xml2txt(javastics_path, workspace_path, usms_list = "banana",
                 dir_per_usm_flag = FALSE, verbose = FALSE)

# In a specific folder
gen_usms_xml2txt(javastics_path, workspace_path, usms_list = "banana", 
                 target_path = output_path, dir_per_usm_flag = FALSE,
                 verbose = FALSE)


```
