---
title: "Generating Stics text files from XML files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating_Stics_text_files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params:
  javastics_path: /tmp/JavaSTICS-v141-stics-v9.1
  workspace_path: /tmp/JavaSTICS-v141-stics-v9.1/example
  output_path: /tmp/example
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup, echo= TRUE, eval = TRUE}
library(SticsRFiles)
library(SticsOnR)
```


```{r, echo = FALSE, eval = TRUE}
javastics_path <- params$javastics_path
workspace_path <- params$workspace_path
output_path <- params$output_path
```

## Introduction

* In this part information are focusing on functionalities related to generation of text input files for the Stics model.

* Generated files will be directly usable for running the model with **SticsOnR** package functions. See the documentation [here](https://sticsrpacks.github.io/SticsOnR).


## Converting JavaStics workspace files
For using the model directly neither using the JavaStics graphical interface nor the `run_javastics` function interface, we provide a function, `gen_usms_xml2txt`, for converting JavaStics XML files to Stics text files from a JavaStics workspace. The model is then run using the [`run_stics()`](https://sticsrpacks.github.io/SticsOnR/reference/run_stics.html) function from [SticsOnR](https://sticsrpacks.github.io/SticsOnR/index.html)

Observation files may also be copied if they have a standard name as an usm name and a `.obs` extension. If not, they must be renamed to do so.

```{r paths_def}
# Specifying the JavaStics folder
javastics_path <- "/path/to/JavaSTICS-1.41-stics-9.1"

# Specifying a workspace as a subfolder of JavaStics 
workspace_path <- "example"
# or an absolute path to an extrenal folder
# workspace_path <- "/path/to/javastics/workspace"

# Specifying an output folder path 
output_path <- "/path/to/output/folder"

```

## Converting files into separated folders (one per usm)
```{r convert_to_separate}
# Generating files for all the usms contained in the workspace
# Into the workspace
gen_usms_xml2txt(javastics_path, workspace_path)
# Into a specific output folder
gen_usms_xml2txt(javastics_path, workspace_path, target_path = output_path)

# Generating files for a subset of usms
# Into the workspace
gen_usms_xml2txt(javastics_path, workspace_path, usms_list = c("banana","wheat"))
# Into a specific folder
gen_usms_xml2txt(javastics_path, workspace_path, usms_list = c("banana","wheat"), target_path = output_path)

# Getting returned information about files generation
gen_info <- gen_usms_xml2txt(javastics_path, workspace_path, usms_list = c("banana","wheat"), target_path = output_path)

gen_info
#> $usms_paths
#> [1] "banana" "wheat" 
#> 
#> $files_path
#>  [1] "/tmp/JavaSTICS-v141-stics-v9.1/example/climat.txt"     
#>  [2] "/tmp/JavaSTICS-v141-stics-v9.1/example/param.sol"      
#>  [3] "/tmp/JavaSTICS-v141-stics-v9.1/example/ficini.txt"     
#>  [4] "/tmp/JavaSTICS-v141-stics-v9.1/example/ficplt1.txt"    
#>  [5] "/tmp/JavaSTICS-v141-stics-v9.1/example/fictec1.txt"    
#>  [6] "/tmp/JavaSTICS-v141-stics-v9.1/example/station.txt"    
#>  [7] "/tmp/JavaSTICS-v141-stics-v9.1/example/new_travail.usm"
#>  [8] "/tmp/JavaSTICS-v141-stics-v9.1/example/tempopar.sti"   
#>  [9] "/tmp/JavaSTICS-v141-stics-v9.1/example/tempoparv6.sti" 
#> [10] "/tmp/JavaSTICS-v141-stics-v9.1/example/ficplt2.txt"    
#> [11] "/tmp/JavaSTICS-v141-stics-v9.1/example/fictec2.txt"    
#> 
#> $copy_status
#> [1] TRUE TRUE
#> 
#> $obs_copy_status
#> [1] TRUE TRUE
```

## Converting files into one folder (overwriting case) 
```{r convert_to_one}
# Generating files directly in the workspace or a specific folder (no usm sub-folder) 
# In this case the model files are overwritten at each gen_usms_xml2txt call !
# In the workspace
gen_usms_xml2txt(javastics_path, workspace_path, usms_list = "banana", dir_per_usm_flag = FALSE)

# In a specific folder
gen_usms_xml2txt(javastics_path, workspace_path, usms_list = "banana", 
                 target_path = output_path, dir_per_usm_flag = FALSE)


```
