---
title: "Manipulating Stics XML files"
output: 
   html_document:
     theme: default
     toc: true
     toc_float:
       collapsed: false
       smooth_scroll: false
     number_sections: true
vignette: >
  %\VignetteIndexEntry{Manipulate_Stics_XML_files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params:
  xml_dir: !r system.file(file.path("extdata","xml","examples","V9.0"), package = "SticsRFiles" )
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SticsRFiles)
#print(params$xml_dir)
options(warn=-1)
```


# Introduction
The purpose of this document is to provide starting information about how to use SticsRFiles functions for extracting data from XML Stics input files or replacing data into them.

Using simple functions use cases

Exploring files content for files sets ...


# Understanding what are parameters in Stics XML files

XML files are self documented files and hierarchically structured. Syntaxic rules are based on what we call nodes (tags ?, similar to html ones) and attributes.


Stics XML files are filled with different kinds of parameters description
structures according to their type/role in the model formalisms.

As we can see in the above XML file extract from a general parameter file:    

* Parameters are grouped according to formalisms (in a `xml <formalisme>...</formalisme> ` block)

* Parameters may be defined by single values
This kind of definition is common for all files types and the parameter name is stored in the `nom` attribute of a `param` XML node. So the first parameter name in the example is `beta` and its value is `1.40000` (inside param tags).

* Option parameters (with or without attached parameters, eventually with sub-option)
The corresponding parameter name for an **option** node is stored in an attribute named `nomParam` and the corresponding value is given in the attribute `choix` of the option node. For example, the value of the parameter `codesymbiose` is `2` in the example.


```xml
<formalisme nom="Water absorption and nitrogen content of the plant">
    <param format="real" max="2.0" min="1.0" nom="beta">1.40000</param>
    <param format="real" max="1.0" min="0.2" nom="lvopt">0.50000</param>
    <param format="real" max="0.07" min="0.0050" nom="rayon">0.02000</param>
    <param format="real" max="0.1" min="0.01" nom="difN">0.04500</param>
    <param format="real" max="3.0" min="0.0" nom="concrr">0.02000</param>
    <param format="real" max="30.0" min="5.0" nom="plNmin">10.00000</param>
    <param format="real" max="50.0" min="0.0" nom="irrlev">20.00000</param>
    <param format="real" max="50.0" min="0.0" nom="QNpltminINN">0.00000</param>
    <option choix="2" nom="Nitrogen fixation by legumes" nomParam="codesymbiose">
      <choix code="1" nom="critical nitrogen"/>
      <choix code="2" nom="nodule activity">
        <option choix="2" nom="mineral nitrogen inhibition" nomParam="codefxn">
          <choix code="1" nom="no effect"/>
          <choix code="2" nom="nitrogen amount"/>
          <choix code="3" nom="nitrogen concentration"/>
        </option>
      </choix>
    </option>
</formalisme>
```

* Parameters with several values (for soil layers, technical operations,...)    
Several values may be defined for example relatively to soil parameters (different layers) or for crop management operations (with time repetitions). 

  * example from soil files    
  In the above file extract for soil parameters, each set of parameters is repeated for each soil layer (up to 5). The parameters names are defined in a `nom` attribute for each `colonne` XML tag.    
  
```xml

<sol nom="solcanne">

...

<tableau nb_colonnes="8" nom="layer 1">
  <colonne nom="epc">20.00</colonne>
  <colonne nom="HCCF">46.80</colonne>
  <colonne nom="HMINF">26.20</colonne>
  <colonne nom="DAF">1.08</colonne>
  <colonne nom="cailloux">0.00</colonne>
  <colonne nom="typecailloux">1</colonne>
  <colonne nom="infil">50.00</colonne>
  <colonne nom="epd">10</colonne>
</tableau>

...

<tableau nb_colonnes="8" nom="layer 5">
<colonne nom="epc">20.00</colonne>
<colonne nom="HCCF">50.10</colonne>
<colonne nom="HMINF">25.50</colonne>
<colonne nom="DAF">0.99</colonne>
<colonne nom="cailloux">0.00</colonne>
<colonne nom="typecailloux">1</colonne>
<colonne nom="infil">50.00</colonne>
<colonne nom="epd">10</colonne>
</tableau>
</sol>

...

```

  * example from tec files    
  In the above file extract for crop management parameters, each set of parameters is repeated for each operation (undefined number). The parameters names are defined in a `nom` attribute for each `colonne` XML tag, inside an `intervention` node.
  
```xml
<formalisme nom="supply of organic residus">
        <ta nb_interventions="5" nom="interventions">
            <ta_entete nb_colonnes="7">
                <colonne nom="julres"/>
                <colonne nom="coderes"/>
                <colonne nom="qres"/>
                <colonne nom="Crespc"/>
                <colonne nom="CsurNres"/>
                <colonne nom="Nminres"/>
                <colonne nom="eaures"/>
            </ta_entete>
            <intervention nb_colonnes="7">
                <colonne nom="julres">112</colonne>
                <colonne nom="coderes">1</colonne>
                <colonne nom="qres">1.00</colonne>
                <colonne nom="Crespc">42.00</colonne>
                <colonne nom="CsurNres">60.00</colonne>
                <colonne nom="Nminres">0.00</colonne>
                <colonne nom="eaures">0.00</colonne>
            </intervention>
            
            ...
            
            <intervention nb_colonnes="7">
                <colonne nom="julres">220</colonne>
                <colonne nom="coderes">2</colonne>
                <colonne nom="qres">1.00</colonne>
                <colonne nom="Crespc">42.00</colonne>
                <colonne nom="CsurNres">60.00</colonne>
                <colonne nom="Nminres">0.00</colonne>
                <colonne nom="eaures">0.00</colonne>
            </intervention>
        </ta>
    </formalisme>
```






# Retrieving parameters information from files

Rules have been defined to search and extract information inside XML files in order to simplify the way of using functions dedicated to XML files manipulations.    

In some cases information relative to upward dependence are needed for extracting parameters values, but in most cases only the parameters names are mandatory in functions arguments.

## Getting parameters names list from a file or a file list    


```{r}
sols <- file.path(params$xml_dir,"sols.xml")

get_files_params_names(sols)
```

```{r}
files_list <- c(file.path(params$xml_dir,"param_gen.xml"),
                file.path(params$xml_dir,"sols.xml"))

param_names <- get_files_params_names(files_list)

head(param_names[[1]])

head(param_names[[2]])


```


## Getting parameters values from a file    
* Scalar parameters
  * for one parameter and several occurence    
```{r}
sols <- file.path(params$xml_dir,"sols.xml")

get_param_xml(sols,param_names_list = "argi", show_xpath = T )
```    
  * For one parameter and one occurence (conditional selection)    
```{r}
sols <- file.path(params$xml_dir,"sols.xml")

get_param_xml(sols,param_names_list = "argi", parent_name = "sol",
                        parent_sel_attr = "solcanne", show_xpath = T)
```    
  * for several parameters and several occurrence    
```{r}
sols <- file.path(params$xml_dir,"sols.xml")

get_param_xml(sols,param_names_list = c("argi", "pH"), show_xpath = T )
```    
  * for several parameters and one occurrence (conditional selection)    
  **THE RESULTS IS FALSE**
```{r}
sols <- file.path(params$xml_dir,"sols.xml")

get_param_xml(sols,param_names_list = c("argi", "pH") , parent_name = "sol",
                        parent_sel_attr = "solcanne", show_xpath = T)
```
```{r}
sols <- file.path(params$xml_dir,"param_gen.xml")

get_param_xml(sols,param_names_list = "iniprofil", show_xpath = T )
```

* Vector parameters
  * All values (for all soil layers, or crop management operations)

  * Selecting   
  
  
## Getting parameters values for a files list 



# Replacing parameters values in files



# Using functions for multiple files



