---
title: "Manipulating Stics XML files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Manipulate_Stics_XML_files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy=TRUE
)
```


```{r}
library(SticsRFiles)
```

```{r setup, results="hide", echo=FALSE}
#options(warn=-1)
xml_root <- normalizePath(tempdir(), winslash = "/", mustWork = FALSE)
xml_loc_dir <- file.path(xml_root,"XML")
if(!dir.exists(xml_loc_dir)) dir.create(xml_loc_dir)

```




```{r rm_files, results='hide', echo=F}
files_list <- list.files(path = xml_loc_dir, pattern = "\\.xml$", full.names = T)
files_list
if (length(files_list))  {
  print("Removing files")
  file.remove(files_list)
}
```


# Introduction
The purpose of this document is to provide starting information about:

* how parameters are defined/structured in Stics XML files
* how to find information about parameters among an XML files set
* how to use functions for extracting data from XML Stics input files
* how to use functions for replacing data into them

Simple reproducible use cases based on 3 essential functions **get_param_info**, **get_param_xml** and **set_param_xml** are described in this document.


For the moment, queries for selecting information among `usms`, `soils` are limited.
For example, getting soil parameters for a list of soils is possible, as getting usms informations using a usms list.

Queries inspecting elements downwards are not available yet; as for example, getting soil names/parameters using a criterion like `argi` values > 20.

Exploring files content for files sets are also possible, but not replacing parameters values.


# Understanding what are parameters in Stics XML files

## XML files structures describing parameters 

XML files are self documented files and hierarchically structured. Syntaxic rules are based on what we call nodes (i.e. nodes names, similar to html tags) and attributes (nodes attributes values).


Stics XML files are filled with different kinds of parameters description
structures according to their type/role in the model formalisms.

As we can see in the above XML file extract from a general parameter file:    

* Parameters are grouped according to formalisms (in a `xml <formalisme>...</formalisme> ` block)

* Parameters may be defined by single values    
This kind of definition is common for all files types and the parameter name is stored in the `nom` attribute of a `param` XML node. So the first parameter name in the example is `beta` and its value is `1.40000` (inside `param` tags).

* Option parameters (with or without attached parameters, eventually with sub-option)    
The corresponding parameter name for an **option** node is stored in an attribute named `nomParam` and the corresponding value is given in the attribute `choix` of the option node. For example, the value of the parameter `codesymbiose` is `2` in the example.


```xml
<formalisme nom="Water absorption and nitrogen content of the plant">
<param format="real" max="2.0" min="1.0" nom="beta">1.40000</param>
<param format="real" max="1.0" min="0.2" nom="lvopt">0.50000</param>
<param format="real" max="0.07" min="0.0050" nom="rayon">0.02000</param>
<param format="real" max="0.1" min="0.01" nom="difN">0.04500</param>
<param format="real" max="3.0" min="0.0" nom="concrr">0.02000</param>
<param format="real" max="30.0" min="5.0" nom="plNmin">10.00000</param>
<param format="real" max="50.0" min="0.0" nom="irrlev">20.00000</param>
<param format="real" max="50.0" min="0.0" nom="QNpltminINN">0.00000</param>
<option choix="2" nom="Nitrogen fixation by legumes" nomParam="codesymbiose">
<choix code="1" nom="critical nitrogen"/>
<choix code="2" nom="nodule activity">
<option choix="2" nom="mineral nitrogen inhibition" nomParam="codefxn">
<choix code="1" nom="no effect"/>
<choix code="2" nom="nitrogen amount"/>
<choix code="3" nom="nitrogen concentration"/>
</option>
</choix>
</option>
</formalisme>
```

* Parameters with several values (for soil layers, technical operations,...)     
Several values may be defined for example relatively to soil parameters (different layers) or for crop management operations (with time repetitions). 

* Example from soil files     
In the above file extract for soil parameters, each set of parameters is repeated for each soil layer (up to 5). The parameters names are defined in a `nom` attribute for each `colonne` XML tag.    

```xml

<sol nom="solcanne">

...

<tableau nb_colonnes="8" nom="layer 1">
<colonne nom="epc">20.00</colonne>
<colonne nom="HCCF">46.80</colonne>
<colonne nom="HMINF">26.20</colonne>
<colonne nom="DAF">1.08</colonne>
<colonne nom="cailloux">0.00</colonne>
<colonne nom="typecailloux">1</colonne>
<colonne nom="infil">50.00</colonne>
<colonne nom="epd">10</colonne>
</tableau>

...

<tableau nb_colonnes="8" nom="layer 5">
<colonne nom="epc">20.00</colonne>
<colonne nom="HCCF">50.10</colonne>
<colonne nom="HMINF">25.50</colonne>
<colonne nom="DAF">0.99</colonne>
<colonne nom="cailloux">0.00</colonne>
<colonne nom="typecailloux">1</colonne>
<colonne nom="infil">50.00</colonne>
<colonne nom="epd">10</colonne>
</tableau>
</sol>

...

```

* Example from tec files    
In the above file extract for crop management parameters, each set of parameters is repeated for each operation (undefined number). The parameters names are defined in a `nom` attribute for each `colonne` XML tag, inside an `intervention` node.

```xml
<formalisme nom="supply of organic residus">
<ta nb_interventions="5" nom="interventions">
<ta_entete nb_colonnes="7">
<colonne nom="julres"/>
<colonne nom="coderes"/>
<colonne nom="qres"/>
<colonne nom="Crespc"/>
<colonne nom="CsurNres"/>
<colonne nom="Nminres"/>
<colonne nom="eaures"/>
</ta_entete>
<intervention nb_colonnes="7">
<colonne nom="julres">112</colonne>
<colonne nom="coderes">1</colonne>
<colonne nom="qres">1.00</colonne>
<colonne nom="Crespc">42.00</colonne>
<colonne nom="CsurNres">60.00</colonne>
<colonne nom="Nminres">0.00</colonne>
<colonne nom="eaures">0.00</colonne>
</intervention>

...

<intervention nb_colonnes="7">
<colonne nom="julres">220</colonne>
<colonne nom="coderes">2</colonne>
<colonne nom="qres">1.00</colonne>
<colonne nom="Crespc">42.00</colonne>
<colonne nom="CsurNres">60.00</colonne>
<colonne nom="Nminres">0.00</colonne>
<colonne nom="eaures">0.00</colonne>
</intervention>
</ta>
</formalisme>
```

## Functions for XML files manipulations
Rules have been defined to easily search and extract information from XML files in order to simplify the way of using functions dedicated to XML files manipulations.

In some cases, information relative to upward dependence are needed for extracting parameters values, but in most cases only the parameters names are mandatory in functions arguments.


# Getting XML files examples from `SticsRFiles` library

Several XML examples files have bee included  in the package in order to use them in `reproducible`
**manipulations** and **results** described in this document. These examples files are extracted from the JavaStics 
standard distribution (some of them have been renamed).

## Getting the XML files source directory 
In xml_dir, we store the directory path of the XML files available in the SticsRFiles installation directory.

```{r set_xml_dir}
xml_dir <- get_examples_path( file_type = "xml", version_name = "V9.1")

# For linux 
#> "/path/to/user/R/x86_64-pc-linux-gnu-library/3.6/SticsRFiles/extdata/xml/examples/V9.1"

# For windows
#> "C:/Users/username/Documents/R/win-lib/3.6/SticsRFiles/extdata/xml/examples/V9.1"

```

## List examples files

```{r}

xml_files <- list.files(path = xml_dir, pattern = ".xml$", full.names = T)

# Listing only the first three files of the entire list 

# head(xml_files, n = 3)
# For linux 
#> [1] "/path/to/user/R/x86_64-pc-linux-gnu-library/3.6/SticsRFiles/extdata/xml/examples/V9.1/file_ini.xml"     
#> [2] "/path/to/user/R/x86_64-pc-linux-gnu-library/3.6/SticsRFiles/extdata/xml/examples/V9.1/file_plt.xml"     
#> [3] "/path/to/user/R/x86_64-pc-linux-gnu-library/3.6/SticsRFiles/extdata/xml/examples/V9.1/file_sta.xml"     

# For windows
#> [1] "C:/Users/username/Documents/R/win-lib/3.6/SticsRFiles/extdata/xml/examples/V9.1/file_ini.xml"   
#> [2] "C:/Users/username/Documents/R/win-lib/3.6/SticsRFiles/extdata/xml/examples/V9.1/file_plt.xml"    
#> [3] "C:/Users/username/Documents/R/win-lib/3.6/SticsRFiles/extdata/xml/examples/V9.1/file_sta.xml"     

```

``` {r, echo=FALSE, results="hide"}
files_list <- list.files(path = xml_dir, pattern = ".xml$")
legend <- c("initializations", "plant", "station", "crop management",
               "general", "general (new formalisms)", "soils", "usms" )           
dt <- data.frame(files = files_list, groups=legend )

```

The correponding files types to file names is given in the above table
``` {r, echo = FALSE}
                        
knitr::kable(dt, caption = "Correspondence between XML example files names and parameters groups")

```


## Copying files in a local directory

### A specific file


```{r}
# Setting a local directory path
# xml_loc_dir <- "/path/to/local/directory"

file.copy(from = file.path(xml_dir,"sols.xml"), to = xml_loc_dir)
```

### All files

```{r}

file.copy(from = xml_files,to = xml_loc_dir)
```


# Getting information about parameters

## Searching parameters names from files

The `get_param_info` function only take names or names substring of parameters or a list of. The search is performed in an XML files list depending on the model version (9.1 as the default one).

In that case, no need to care of which XML file(s) may contain the searched parameter(s). 

Some warnings may be displayed in case of duplicated bounds values. In that case retained values are indicated in warning messages.

The information columns in the returned tibble are:

* `name`: parameter name
* `file`: XML file name
* `min`: minimal bound
* `max`: maximal bound
* `formalism`: formalism short description

Whatever the arguments list is when calling the function, the returned tibble always contains these columns.


Information about bounds or formalism, may be missing according to the parameter kind (initialization, usm), or type (character).


### Getting all STICS parameters information


Using the `get_param_info` function whithout any argument, or with specifying the desired STICS version, allows to get a table containing all the input parameters.

```{r}

param_names <- get_param_info()

# Displaying the returned data as a paged table
rmarkdown::paged_table(param_names)

```


### Finding one or more parameters names

```{r, results="markup"}

get_param_info(parameter = "albedo")

get_param_info(parameter = c("albedo", "latitude", "humcapil"))

```

### Finding parameters names using partial names

```{r}

get_param_info( parameter = "hum")


param_names <- get_param_info(parameter = c("al","hum"))

# Displaying the returned data as a paged table
rmarkdown::paged_table(param_names)

```


## Searching parameters attached to formalisms

### Finding formalisms using partial names (i.e. short description)

```{r, warning=FALSE}

get_param_info(formalism = "yield")


param_formalisms <- get_param_info(formalism = c("yield","leaves"))

head(param_formalisms)

```


### Finding formalisms using partial formalism and parameters names

```{r, warning=FALSE}

get_param_info(formalism = "yield", parameter = "grain")


param_formalisms <- get_param_info(formalism = "yield", parameter = c("inflo","lai"))

rmarkdown::paged_table(param_formalisms)

```



## Searching information using keywords

The keyword argument (one or several strings) of the function may be used to search in all textual columns as `name`, `file` or `formalism`.  


```{r, warning=FALSE}

get_param_info(keyword = "plant")


param_names <- get_param_info(keyword = c("plant","leave"))

# Displaying the returned data as a paged table
rmarkdown::paged_table(param_names)

```




# Getting parameters values from a file

The `get_param_xml` function is used for extracting parameters values in XML files, providing it a parameter name or a names list.

For usms or soils parameter, a conditional selection may be used to filter parameter values respectively through a list of usms or soils. 


## Scalar parameters
* For one parameter and one occurence    
```{r}
# Fixing files paths
sols <- file.path(xml_loc_dir,"sols.xml")
par_gen <- file.path(xml_loc_dir,"param_gen.xml")

# A option parameter
get_param_xml(par_gen, param_name = "codeactimulch" )

# A simple parameter
get_param_xml(par_gen, param_name = "tnitopt" )

# Using a conditional selection
get_param_xml(sols,param_name = "argi", select = "sol",
              value = "solcanne")

```

* For one parameter and several occurences    
```{r}

# For all soils
get_param_xml(sols,param_name = "argi" )
```    
  
* For several parameters and several occurrences    
```{r}

# For all soils
get_param_xml(sols,param_name = c("argi", "pH") )
```    
* For several parameters and one occurrence (conditional selection)    

```{r}

# For one soil
get_param_xml(sols,param_name = c("argi", "pH") , select = "sol",
              value = "solcanne")
```


## Vector parameters
* All values (for all soil layers, or crop management operations)
```{r}
# For all soil layers
get_param_xml(sols, param_name = c("epc", "infil"), select = "sol",
              value = "solcanne" )
```

```{r}
# For all irrigation operations
tec <- file.path(xml_loc_dir,"file_tec.xml")
get_param_xml(tec, param_name = c("julapI_or_sum_upvt", "amount"))

```

* Selecting with values index   

```{r}
# For soil layers 1 to 3
get_param_xml(sols,param_name = c("epc", "infil"), select = "sol", value = "solcanne", ids = 1:3 )
```

```{r}

# For irrigation operations 1 to 5
get_param_xml(tec, param_name = c("julapI_or_sum_upvt", "amount"), ids = 1:5)

```


## Getting all parameters values

Using the function whithout any argument produces a list of all parameters values from a file.

```{r}
tec_param_values <- get_param_xml(tec)[[1]]
# 
# Displaying only a subset of the list
head(tec_param_values, n = 10)

```

A files list may be used also to get all parameters as follows in the same list


```{r}
param_values <- get_param_xml(c(tec,sols))

# Sub-list names
names(param_values)

# param_values extract for each sub-list
head(param_values$file_tec.xml, n = 5)

head(param_values$sols.xml, n = 5)
```


# Replacing parameters values in files

## Defining the output file from functions arguments
The function **set_param_xml** allows to set the output file with the following optional arguments as follows:

* For overwriting the existing file: use ```overwrite = TRUE```

* For generating a file in another directory: use `out_path` for setting an output directory

* Specifying a new file name: use  `out_file` for giving the new file name (inculding the .xml extension)



## Scalar parameters
* For one parameter and one occurence    
```{r}


# A option parameter
set_param_xml(par_gen,param_name = "codeactimulch", param_value = 2, overwrite = T )

# Controlling written values
get_param_xml(par_gen, param_name = "codeactimulch" )

# A simple parameter
set_param_xml(par_gen, param_name = "tnitopt", param_value = 29.5, overwrite = T )

# Controlling written values:
get_param_xml(par_gen, param_name = "tnitopt" )

# Using a conditional selection
set_param_xml(sols, param_name = "argi", param_value = 33, select = "sol",
              value = "solcanne", overwrite = T )

# Controlling written values
get_param_xml(sols,param_name = "argi", select = "sol",
              value = "solcanne")

```


* For one parameter and several occurences    
```{r}

# For all soils
# One value per occurence
set_param_xml(sols, param_name = "argi", param_value = list(1:33), overwrite = T )
# Controlling written values
get_param_xml(sols,param_name = "argi" )

# Same value for all occurences
set_param_xml(sols, param_name = "argi", param_value = 40, overwrite = T )

# Controlling written values
get_param_xml(sols,param_name = "argi" )


```    

* For several parameters and several occurrences    
```{r}

# For all soils
# One value per occurence
set_param_xml(sols, param_name = list("argi", "pH"), param_value = list(1:33,33:1), overwrite = T )
# Controlling written values
get_param_xml(sols,param_name = c("argi", "pH") )

# Same value for all occurences
set_param_xml(sols, param_name = c("argi", "pH"), param_value = list(50, 8), overwrite = T )
# Controlling written values
get_param_xml(sols,param_name = c("argi", "pH") )


```    
* For several parameters and one occurrence (conditional selection)    

```{r}

# For one soil
set_param_xml(sols, param_name = c("argi", "pH"), param_value = list(50,8), select = "sol",
              value = "solcanne", overwrite = T )
# Controlling written values
get_param_xml(sols,param_name = c("argi", "pH") , select = "sol",
              value = "solcanne")
```


## Vector parameters
* All values (for all soil layers, or crop management operations)
```{r}
# For all soil layers
set_param_xml(sols, param_name = c("epc", "infil"), param_value = list(18:22,48:52), select = "sol", value = "solcanne", overwrite = T )
# Controlling written values
get_param_xml(sols, param_name = c("epc", "infil"), select = "sol",
              value = "solcanne" )
```

```{r}
# For all irrigation operations
tec <- file.path(xml_loc_dir,"file_tec.xml")
set_param_xml(tec, param_name = c("julapI_or_sum_upvt", "amount"), param_value = list(200:215,20:35), overwrite = T )
# Controlling written values
get_param_xml(tec, param_name = c("julapI_or_sum_upvt", "amount"))

```

* Some values   

```{r}
# For soil layers 1 to 3
set_param_xml(sols, param_name = c("epc", "infil"), param_value = list(20:18,50:48), select = "sol", value = "solcanne", overwrite = T, ids = 1:3 )
# Controlling written values
get_param_xml(sols,param_name = c("epc", "infil"), select = "sol",
              value = "solcanne", ids = 1:3 )
```

```{r}

# For irrigation operations 1 to 5 (same indices for all parameters)
set_param_xml(tec, param_name = c("julapI_or_sum_upvt", "amount"), param_value = list(204:200,24:20), overwrite = T, ids = 1:5)
# Controlling written values
get_param_xml(tec, param_name = c("julapI_or_sum_upvt", "amount"), ids = 1:5)

```








