---
title: "Starting user guide draft and notes for the SticsOnR package"
author: "Patrice Lecharpentier"
date: "`r Sys.Date()`"
#output: html_document
output: rmarkdown::html_vignette
   #toc_float: TRUE 
vignette: >
  %\VignetteIndexEntry{Starting user guide draft and notes for the SticsOnR package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Content draft of the starting guide

### Introduction

* In this part informations are focusing only the functionnalities related to Stics xml files manipulation and generation (mailing capabilities)

* The generation of wheather files and observations files are not included yet in the package


### Content
#### Installation    
  
Before using SticsOnR some dependencies packages must be installed.
     

* CRAN packages (see SticsOnR package DESCRIPTION file)

  * XML (needed by xmlDocument class)
  * dplyr (for manipulating data.frames, tibbles)
  * readxl ( loading XL sheets of inputs parameters )
  * stringr (character strings manipulation)
  * lubridate ( used by other functions for managing stics output data)

* specific package 
     
  * Classes ( providing xmlDocument class for managing xml data )
    The source tar.gz form of the package can be installed from the console    
    
>   install.packages("/path/to/Classes_1.1.tar.gz", repos = NULL, type = "source") 

   or from the **Packages** tab in RStudio selecting **Install from Package archive file (.tar.gz)** in the list


#### Package functions

* Package help
  * in the **Packages** tab in R studio by clicking on package name, access to functions help
  * in the console : ?function_name


* Main functionnalities
      
  * Load an xml file as an xmlDocument object
    * xmldocument
      
  * Get or set parameters values in xmlDocuments
    * get_param_value
    * get_params_from_doc
    * get_params_number
    * set_param_value
      
  * Save an xmlDocument as an xml file
    saveXmlDoc
      

  * Generate and changes parameters values in xmlDocument objects using a table of parameters
    * gen_ini_doc
    * gen_sta_doc
    * gen_tec_doc
    * gen_usms_doc
    * gen_sols_doc

      
  * Generate xml files using a table of parameters     
    * gen_ini_file
    * gen_sta_file
    * gen_tec_file
    * gen_usms_file
    * gen_sols_file

      
  * manipulation of xl sheets
    * get_params_from_table


      
#### Function calling syntax

Take care about using optional arguments with default values. In order to be sure to set a value to the right argument (overloading the default one) the best way is to use the named arguments syntax.


```{r eval=FALSE}  
# Function signature
func <- function(arg1, arg2 = "value2", arg3 = "value3") {}

# Right syntax 
func(arg1, arg3 = "value3")

#Bad syntax, the "value" is assigned to the 2nd argument of the function
func(arg1,"value3")

```



#### Getting example files and scripts included in the package

* Getting inputs XL example file: 

```{r eval=FALSE}
library(SticsOnR)
# getting the file in the current directory
get_inputs_example()
# getting the file in a specic directory
get_inputs_example(dest_dir = "/home/plecharpent/Work/workspace")

> [1] "inputs_stics_example.xlsx  has been copied in directory  
 /home/plecharpent/Work/workspace"

```

* Getting example scripts for producing ini, sols, sta, tec and usms from the package
  * Getting scripts files list     
```{r}
library(SticsOnR)
get_script()

```      
  * Getting a script or all scripts
  
```{r, eval=FALSE}
# a specific
get_script("generate_ini_files.R")
get_script("generate_ini_files.R", dest_dir = "/path/to/dir")

# all scripts
get_script("all")
get_script("all", dest_dir = "/path/to/dir")
```
   * Only adapt the paths / directories and files names. In these scripts the XL file is given just as an example. It will be replaced by by the user ones.
   
   * The table loaded from an XL sheet is entirely used in the examples, but a selection may be made to generate files only for a subset
   
#### List of constraints or limitations
  
* For `usms` and `sols` files: 
       
  * For the moment the functions are not taking into account the xml files content update: i.e. parameters values modification for existing `usm` or `sol`, or adding new usm or sol. This means that xml sols or usms files are newly generated upon values contained in the xl sheets. 
  * So, all the **usm** or **sol** parameters must be present in the XL sheet
    
* For `tec`, `sta` and `ini` files : 
    
  * Take care with consistency of the base xml file provided to be duplicated and modified with parameters present in XL sheets. Messages will be displayed in case of a misspelled parameter, or an inexistent parameter in the xml file which may correspond to another Stics version than the desired one.
  
  * All the parameters are not mandatory in the XL sheet, only those we want to modify



* Note on parameters names syntax
  * The exact parameters names existing in xml files (with respect of case sensitive) must be used generally in XL sheets. But we have added some **shortcuts parameters** names for some of the which are too long or meaningless.
  * For long parameters names or meaningless a names correspondence list have been added to facilitate xl sheet filling. It have been included while retreiving data from data frame with names replacement for existing ones in the correspondence list.     
  
```{r echo=FALSE,warning=FALSE, message=FALSE}
library(knitr)
library(SticsOnR)
dict <- get_params_from_table()
dt <- data.frame(shortcuts=names(dict), real_name=unlist(dict,use.names = FALSE))

kable(dt, caption = "List of possible builtin parameters names substitution")
```

To get this correspondance list : call get_parameters_from_table(), as follows

```{r}
library(knitr)
library(SticsOnR)
get_params_from_table()

```
   * But, it is possible to create and provide a personal list through the `dict` argument of the function.
This argument is usable at the moment for producing only *tec* files.  When calling `gen_tec_file` the argument dict may be set as follows
```{r eval=FALSE}  
gen_tec_file(..., dict = myDict) 
```


* Files names or usms/sols names in xl sheet
  * key words for detecting names or file names column: usm_, tec_, ini_, soil_, sta_

  * files names : may be either full files names or only file names without any suffix like _tec.xml, _ini.xml or _sta.xml 


* Consistency data checks in XL sheets    
For the moment *no consistency checks* functionnalities have been included in the package. So, we must take care of links between xml files names for example files names defined for usms related to names declared in `Tec`, `Station`, and so on. 


* XL sheets names    
XL sheet names are completely free, but must be meaningfull regarding to the kind of parameters they contain.



## Further works

### List of improvements

  * management of xml files : how to organize a set of files common to all Stics versions and specific files to stics versions


### List of functionnalities not included but, guessing they will be useful

  * add a name or names vector as input arg for all functions gen_*_doc and gen_*_file giving the possibility to produce xml doc or files for a subset of the input table (i.e. for a list of usms names)

  * add a functionnality for checking consistency between data contained in XL sheets of xl files of Stics input parameters

  * add functionnalities to generate observations and wheather files from XL sheets

  * add the possibility to use csv separate files instead of xl sheets (and also consistency checks between them !)
  
  * add a function to store in inst folder in examples and generate templates files for xml files from given files (files templates and operations templates) 
  
  * add a function to extract parameters values from an xml file, for a list of parameters names ( get_params_from_xml)


### List of checks to do
  * Verify the updating use case of data in xml files for all functions gen_*_doc !

  * Make a list of possible unit tests and other tests to write for functions



