---
title: "Upgrading STICS XML files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Upgrading STICS XML files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = TRUE
)
```

```{r eval=FALSE, ,include=FALSE}
library(SticsRFiles)
```

```{r include=FALSE}
suppressWarnings(library(SticsRFiles))
```

```{r echo=FALSE}
stics_version <- get_stics_versions_compat()$latest_version
```


The XML files structure has changed either between STICS version **9.2** and versions **10.X.Y** or between STICS versions **10.X.Y** and version 11.0.0 and a transformation is mandatory for upgrading from a major version to a upper one.    

So, to avoid difficulties and errors with manual XML files modifications, functionalities have been developed to do so automatically.

One can either transform each kind of file individually (i.e. _tec, _ini,.. files) or a full workspace directory at once.

Here are some examples of use of upgrade generic functions either from version 9 to 10 or 10 to 11.
The initial version is detected and the target version as well. One have to chain upgrades for converting files version 9 to version 11.



```{r, include = FALSE, eval=FALSE}
example_data <- SticsRFiles::download_data(
  out_dir = tempdir(),
  example_dirs = "study_case_1", "V9.2"
)
workspace <- file.path(example_data, "XmlFiles")
out_dir <- file.path(tempdir(), "XmlFiles_V10.2.0")
if (!dir.exists(out_dir)) dir.create(out_dir)
javastics <-
  "/home/plecharpent/tmp/TEST_UPDATE_XML_V9_V10/JavaSTICS-1.41-stics-9.2"
upgrade_workspace_xml(
  workspace = workspace, javastics = javastics,
  out_dir = out_dir, overwrite = TRUE
)
```

## Upgrading a whole workspace

In that case, a workspace adapted to the new version is recreated.

So, apart from the XML files, some other files as outputs definition files (*.mod), weather files attached to USMs and also observation files (named with the USM name, \*.obs) are copied to the new upgraded workspace.

Take care that, if the JavaStics folder is needed, it must corresponds to the initial STICS version of the workspace you want to upgrade. 

The output directory will contain all needed files to run simulations. It will contain a `plant` sub-directory with all needed plant files and general parameters files.

### Case 1: the initial workspace does not contain all needed files
In that case the JavaStics folder path have to be set in the function inputs
```{r, eval = FALSE, results='markup'}
workspace <- "/path/to/workspace/dir/V9.2"
out_dir <- "/path/to/out/dir/V10"
javastics <- "/path/to/JavaSTICS-1.41-stics-9.2"
upgrade_workspace_xml(
  workspace = workspace,
  javastics = javastics,
  out_dir = out_dir
)
```
The plant files (referenced in the usms.xml file) are not converted and copied in the output directory. Those given
with the JavaStics distribution of the target release have to be used.

### Case 2: the initial workspace contains all the input files
In that case, a sub-directory `plant` exists and contains all the needed plant files (referenced in the usms.xml file).
And the general parameter files also are in the workspace (param_gen.xml and param_newform.xml).
So, no need to specify the JavaStics folder path.

```{r, eval = FALSE, results='markup'}
workspace <- "/path/to/workspace/dir/V9.2"
out_dir <- "/path/to/out/dir/V10"
upgrade_workspace_xml(
  workspace = workspace,
  out_dir = out_dir,
  verbose = TRUE
)
```
Here is the output display indicating the treated files (can be deactivated using the `verbose` argument set to `TRUE`):

<pre style="font-size:11px;">
Upgrading files from version 9 to 10 
From:  /path/to/workspace/dir/V9.2
To:  /path/to/out/dir/V10 
-----------------------------------

param_gen.xml

param_new_form.xml

usms.xml

sols.xml

*_sta.xml

*_ini.xml

*_tec.xml

Copying output definition (*.mod) files.
[1] "/path/to/out/dir/V10/prof.mod" "/path/to/out/dir/V10/rap.mod"  "/path/to/out/dir/V10/var.mod" 
Copying observation (*.obs) files.
 [1] "/path/to/out/dir/V10/banana.obs"             "/path/to/out/dir/V10/BareSoil.obs"          
 [3] "/path/to/out/dir/V10/cc_BristleOat.obs"      "/path/to/out/dir/V10/cc_CrimsonClover.obs"  
 [5] "/path/to/out/dir/V10/cc_ItalianRyegrass.obs" "/path/to/out/dir/V10/cc_mustard.obs"        
 [7] "/path/to/out/dir/V10/cc_vetch.obs"           "/path/to/out/dir/V10/DurumWheat.obs"        
 [9] "/path/to/out/dir/V10/fescue.obs"             "/path/to/out/dir/V10/flax.obs"              
[11] "/path/to/out/dir/V10/grass.obs"              "/path/to/out/dir/V10/lettuce.obs"           
[13] "/path/to/out/dir/V10/maize.obs"              "/path/to/out/dir/V10/potato.obs"            
[15] "/path/to/out/dir/V10/proto_rice.obs"         "/path/to/out/dir/V10/rapeseed.obs"          
[17] "/path/to/out/dir/V10/sorghum.obs"            "/path/to/out/dir/V10/soybean.obs"           
[19] "/path/to/out/dir/V10/strawberry.obs"         "/path/to/out/dir/V10/sugarbeet.obs"         
[21] "/path/to/out/dir/V10/SugarCane.obs"          "/path/to/out/dir/V10/sunflower.obs"         
[23] "/path/to/out/dir/V10/timothy.obs"            "/path/to/out/dir/V10/tomato.obs"            
[25] "/path/to/out/dir/V10/Turmeric.obs"           "/path/to/out/dir/V10/usms_fobs.xml"         
[27] "/path/to/out/dir/V10/vine.obs"               "/path/to/out/dir/V10/wheat.obs"             
Copying LAI dynamics (*.lai) files.
[1] "/path/to/out/dir/V10/ble.lai"     "/path/to/out/dir/V10/bledur.lai" 
[3] "/path/to/out/dir/V10/colza.lai"   "/path/to/out/dir/V10/mais.lai"   
[5] "/path/to/out/dir/V10/salade.lai"  "/path/to/out/dir/V10/soja.lai"   
[7] "/path/to/out/dir/V10/sortie.lai"  "/path/to/out/dir/V10/toursol.lai"
[9] "/path/to/out/dir/V10/vigne.lai"  
Copying weather data (*.YYYY) files.
 [1] "/path/to/out/dir/V10/Auzeville.2012" "/path/to/out/dir/V10/Auzeville.2013"
 [3] "/path/to/out/dir/V10/Auzeville.2014" "/path/to/out/dir/V10/chapaisj.2001" 
 [5] "/path/to/out/dir/V10/clim_snow.2009" "/path/to/out/dir/V10/clim_snow.2010"
 [7] "/path/to/out/dir/V10/climaisj.1996"  "/path/to/out/dir/V10/climbanj.1996" 
 [9] "/path/to/out/dir/V10/climbetj.1994"  "/path/to/out/dir/V10/climbldj.2000" 
[11] "/path/to/out/dir/V10/climbldj.2001"  "/path/to/out/dir/V10/climblej.1994" 
[13] "/path/to/out/dir/V10/climblej.1995"  "/path/to/out/dir/V10/climcanj.1998" 
[15] "/path/to/out/dir/V10/climcanj.1999"  "/path/to/out/dir/V10/climcolj.1994" 
[17] "/path/to/out/dir/V10/climcolj.1995"  "/path/to/out/dir/V10/climfetj.2004" 
[19] "/path/to/out/dir/V10/climfraj.2001"  "/path/to/out/dir/V10/climlinj.2002" 
[21] "/path/to/out/dir/V10/climpdtj.1997"  "/path/to/out/dir/V10/climsalj.2000" 
[23] "/path/to/out/dir/V10/climsnuj.1998"  "/path/to/out/dir/V10/climsojj.1990" 
[25] "/path/to/out/dir/V10/climsorj.1996"  "/path/to/out/dir/V10/climtomj.2000" 
[27] "/path/to/out/dir/V10/climtomj.2001"  "/path/to/out/dir/V10/climtouj.1996" 
[29] "/path/to/out/dir/V10/climvigj.2003"  "/path/to/out/dir/V10/climvigj.2004" 
[31] "/path/to/out/dir/V10/flux000j.2013"  "/path/to/out/dir/V10/fourquej.1988" 
[33] "/path/to/out/dir/V10/prairiej.1995"  "/path/to/out/dir/V10/readingj.2004" 
[35] "/path/to/out/dir/V10/statio5j.1992"  "/path/to/out/dir/V10/statio5j.1993" 
[37] "/path/to/out/dir/V10/statio5j.1994"  "/path/to/out/dir/V10/statio5j.1995" 
[39] "/path/to/out/dir/V10/statio5j.1996" 
-----------------------------------
Files upgrade and copy is complete.
</pre>



## Upgrading a specific file kind

For each kind of file, a specific function is to be used for upgrading its format.

* **`usms.xml`** : upgrade_usms_xml
* **`*_ini.xml`** files: upgrade_ini_xml
* **`param_gen.xml`**: upgrade_param_gen_xml
* **`param_newform.xml`**: upgrade_param_newform_xml
* **`*_plt.xml`** files: upgrade_plt_xml
* **`sols.xml`**: upgrade_sols_xml
* **`*_sta.xml`** files: upgrade_sta_xml
* **`*_tec.xml`** files: upgrade_tec_xml

The input arguments of these functions may vary, refer to each function help for details
about its use.    

Sometimes, general parameter files (`param_gen.xml` or/and `param_newform.xml`) corresponding to the initial version are needed as function inputs because of parameters or options codes transfer to specific files. 

For example, the **`usms.xml`** file is to be upgraded as follows:

```{r, include = FALSE, eval=FALSE}
usms_path <- file.path(example_data, "XmlFiles", "usms.xml")
out_dir <- file.path(tempdir(), "XmlFiles_V10.2.0")
param_gen_path <- file.path(javastics, "config", "param_gen.xml")
upgrade_usms_xml(
  file = usms_path,
  param_gen_file = param_gen_path,
  out_dir = out_dir, overwrite = TRUE
)
```

```{r, eval = FALSE}
usms_path <- "/path/to/workspace/dir/V9.2/usms.xml"
out_dir <- "/path/to/workspace/dir/V10"
param_gen_path <- "/path/to/JavaSTICS-1.41-stics-9.2/config/param_gen.xml"
# or param_gen_path <- "/path/to/workspace/param_gen.xml"
upgrade_usms_xml(
  file = usms_path,
  param_gen_file = param_gen_path,
  out_dir = out_dir
)
```



## Warnings and messages

When **new parameters** are introduced in an upper version, when converting them, their values can be fixed
with default values. But, if their values depend on other parameters values, specific values 
will be used.

For example, when parameters values depend on species, for some of them, parameters values may exist
and will be used.
But if they are not, a message will be displayed indicating that default values are
assigned to parameters.






