---
title: "Upgrading STICS XML files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Upgrading STICS XML files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = TRUE
)
```

```{r eval=FALSE}
library(SticsRFiles)
```

```{r include=FALSE}
suppressWarnings(library(SticsRFiles))
```

```{r echo=FALSE}
stics_version <- get_stics_versions_compat()$latest_version
```


The XML files structure has changed either between STICS version **9.2** and versions **10.X.Y** or between STICS versions **10.X.Y** and version 11.0.0 and a transformation is mandatory for upgrading from a major version to a upper one.    

So, to avoid difficulties and errors with manual XML files modifications, functionalities have been developed to do so automatically.

One can either transform each kind of file individually (i.e. _tec, _ini,.. files) or a full workspace directory at once.

Here are some examples of use of upgrade generic functions either from version 9 to 10 or 10 to 11.



```{r, include = FALSE, eval=FALSE}
example_data <- SticsRFiles::download_data(
  out_dir = tempdir(),
  example_dirs = "study_case_1", "V9.2"
)
workspace <- file.path(example_data, "XmlFiles")
out_dir <- file.path(tempdir(), "XmlFiles_V10.2.0")
if (!dir.exists(out_dir)) dir.create(out_dir)
javastics <-
  "/home/plecharpent/tmp/TEST_UPDATE_XML_V9_V10/JavaSTICS-1.41-stics-9.2"
upgrade_workspace_xml(
  workspace = workspace, javastics = javastics,
  out_dir = out_dir, overwrite = TRUE
)
```

## Upgrading a whole workspace

In that case, a workspace adapted to the new version is recreated.

So, apart from the XML files, some other files as outputs definition files (*.mod), weather files attached to USMs and also observation files (named with the USM name, \*.obs) are copied to the new upgraded workspace.

Take care that, if the JavaStics folder is needed, it must corresponds to the initial STICS version of the workspace you want to upgrade. 

### Case 1: the initial workspace does not contain all needed files
In that case the JavaStics folder path have to be set in the function inputs
```{r, eval = FALSE, results='markup'}
workspace <- "/path/to/workspace/dir/V9.2"
out_dir <- "/path/to/out/dir/V10.2.0"
javastics <- "/path/to/JavaSTICS-1.41-stics-9.2"
upgrade_workspace_xml(
  workspace = workspace,
  javastics = javastics,
  out_dir = out_dir
)
```
### Case 2: the initial workspace contains all the input files
In that case no need to specify the JavaStics folder path
```{r, eval = FALSE, results='markup'}
workspace <- "/path/to/workspace/dir/V9.2"
out_dir <- "/path/to/out/dir/V10.2.0"
upgrade_workspace_xml(
  workspace = workspace,
  out_dir = out_dir
)
```
Here is the output display indicating the treated files (can be deactivated using the `verbose` argument):

<pre style="font-size:11px;">
Upgrading files from version V9.2 to V10.2.0    
From:  /path/to/workspace/dir/V9.2   
To:  /path/to/out/dir/V10.2.0    
-----------------------------------   
param_gen.xml   
param_new_form.xml   
usms.xml   
sols.xml   
*_sta.xml   
*_ini.xml   
*_tec.xml   
Copying *.mod files.   
Copying *.obs files.   
Copying weather files.   
*_plt.xml    
-----------------------------------   
Files upgrade and copy is complete.   
</pre>



## Upgrading a specific file kind

For each kind of file, a specific function is to be used for upgrading its format.

* **`usms.xml`** : upgrade_usms_xml
* **`*_ini.xml`** files: upgrade_ini_xml
* **`param_gen.xml`**: upgrade_param_gen_xml
* **`param_newform.xml`**: upgrade_param_newform_xml
* **`*_plt.xml`** files: upgrade_plt_xml
* **`sols.xml`**: upgrade_sols_xml
* **`*_sta.xml`** files: upgrade_sta_xml
* **`*_tec.xml`** files: upgrade_tec_xml

The input arguments of these functions may vary, refer to each function help for details
about its use.    

Sometimes, general parameter files (`param_gen.xml` or/and `param_newform.xml`) corresponding to the initial version are needed as function inputs because of parameters or options codes transfer to specific files. 

For example, the **`usms.xml`** file is to be upgraded as follows:

```{r, include = FALSE, eval=FALSE}
usms_path <- file.path(example_data, "XmlFiles", "usms.xml")
out_dir <- file.path(tempdir(), "XmlFiles_V10.2.0")
param_gen_path <- file.path(javastics, "config", "param_gen.xml")
upgrade_usms_xml(
  file = usms_path,
  param_gen_file = param_gen_path,
  out_dir = out_dir, overwrite = TRUE
)
```

```{r, eval = FALSE}
usms_path <- "/path/to/workspace/dir/V9.2/usms.xml"
out_dir <- "/path/to/workspace/dir/V10.2.0"
param_gen_path <- "/path/to/JavaSTICS-1.41-stics-9.2/config/param_gen.xml"
# or param_gen_path <- "/path/to/workspace/param_gen.xml"
upgrade_usms_xml(
  file = usms_path,
  param_gen_file = param_gen_path,
  out_dir = out_dir
)
```








