---
title: "Upgrading STICS XML files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Upgrading STICS XML files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = TRUE
)
```

```{r eval=FALSE}
library(SticsRFiles)
```

```{r include=FALSE}
suppressWarnings(library(SticsRFiles))
```

```{r echo=FALSE}
stics_version <- get_stics_versions_compat()$latest_version
```


The XML files structure has changed either between STICS version **9.2** and versions **10.X.Y** or between STICS versions **10.X.Y** and version 11.0.0 and a transformation is mandatory for upgrading from a major version to a upper one.    

So, to avoid difficulties and errors with manual XML files modifications, functionalities have been developed to do so automatically.

One can either transform a complete workspace directory at once or each kind of file individually.

Here some examples of use of upgrade functions either for a whole workspace or for specific files (i.e. _tec, _ini,.. files) from version 9 to 10.

For upgrading from STICS version 10 to 11, just change the functions suffix from _9_10 to _10_11

```{r, include = FALSE, eval=FALSE}
example_data <- SticsRFiles::download_data(
  out_dir = tempdir(),
  example_dirs = "study_case_1", "V9.2"
)
workspace <- file.path(example_data, "XmlFiles")
out_dir <- file.path(tempdir(), "XmlFiles_V10.2.0")
if (!dir.exists(out_dir)) dir.create(out_dir)
javastics <-
  "/home/plecharpent/tmp/TEST_UPDATE_XML_V9_V10/JavaSTICS-1.41-stics-9.2"
upgrade_workspace_xml_9_10(
  workspace = workspace, javastics = javastics,
  out_dir = out_dir, overwrite = TRUE
)
```

## Upgrading a whole workspace

In that case, a workspace adapted to the new version is recreated.

So, apart from the XML files, some other files as outputs definition files (*.mod), weather files attached to USMs and also observation files (named with the USM name, \*.obs) are copied to the new upgraded workspace.

Take care that the JavaStics folder corresponds to the initial STICS version of the workspace you want to upgrade. 

```{r, eval = FALSE, results='markup'}
workspace <- "/path/to/workspace/dir/V9.2"
out_dir <- "/path/to/out/dir/V10.2.0"
javastics <- "/path/to/JavaSTICS-1.41-stics-9.2"
upgrade_workspace_xml_9_10(
  workspace = workspace, javastics = javastics,
  out_dir = out_dir
)
```
Here is the output display indicating the treated files:

<pre style="font-size:11px;">
Upgrading files from version V9.2 to V10.2.0    
From:  /path/to/workspace/dir/V9.2   
To:  /path/to/out/dir/V10.2.0    
-----------------------------------   
param_gen.xml   
param_new_form.xml   
usms.xml   
sols.xml   
*_sta.xml   
*_ini.xml   
*_tec.xml   
Copying *.mod files.   
Copying *.obs files.   
Copying weather files.   
*_plt.xml    
-----------------------------------   
Files upgrade and copy is complete.   
</pre>



## Upgrading a specific file kind

For each kind of file, a specific function is to be used for upgrading its format.

* **`usms.xml`** : upgrade_usms_xml_9_10
* **`*_ini.xml`** files: upgrade_ini_xml_9_10
* **`param_gen.xml`**: upgrade_param_gen_xml_9_10
* **`param_newform.xml`**: upgrade_param_newform_xml_9_10
* **`*_plt.xml`** files: upgrade_plt_xml_9_10
* **`sols.xml`**: upgrade_sols_xml_9_10
* **`*_sta.xml`** files: upgrade_sta_xml_9_10
* **`*_tec.xml`** files: upgrade_tec_xml_9_10

The input arguments of these functions may vary, refer to each function help for details
about its use.    

Sometimes, general parameter files are needed as function inputs because of parameters or options codes movement to specific files. 

For example, the **`usms.xml`** file is to be upgraded as follows:

```{r, include = FALSE, eval=FALSE}
usms_path <- file.path(example_data, "XmlFiles", "usms.xml")
out_dir <- file.path(tempdir(), "XmlFiles_V10.2.0")
param_gen_path <- file.path(javastics, "config", "param_gen.xml")
upgrade_usms_xml_9_10(
  file = usms_path, param_gen_file = param_gen_path,
  out_dir = out_dir, overwrite = TRUE
)
```

```{r, eval = FALSE}
usms_path <- "/path/to/workspace/dir/V9.2/usms.xml"
out_dir <- "/path/to/workspace/dir/V10.2.0"
param_gen_path <- "/path/to/JavaSTICS-1.41-stics-9.2/config/param_gen.xml"
upgrade_usms_xml_9_10(
  file = usms_path, param_gen_file = param_gen_path,
  out_dir = out_dir
)
```








