---
title: "SticsRFiles: tutorial for manipulating XML files and generating model text files"
output: 
  learnr::tutorial:
    theme: cerulean
    #theme: united
    ace_theme: merbivore

# output:
#   slidy_presentation:
#     font_adjustment: -2
#     footer: "SticsRPacks Team, 2020."
runtime: shiny_prerendered
params:
  gen_help: TRUE
  javastics: JavaSTICS-1.41-stics-9.1
  workspace_xml: JavaSTICS-1.41-stics-9.1/example
  workspace_txt: example
  use_default: TRUE
  work_dir: !r file.path(Sys.getenv("HOME"),"R","SticsRfiles-tuto-xml-files")
  #alt_dir: !r file.path(tempdir(), "get-xml-infos")
---

```{r set_var, echo=FALSE, results='hide'}
#library(SticsRFiles)

work_dir <- params$work_dir

# Checking OS
if (Sys.info()['sysname'] == "Windows") work_dir <- shortPathName(work_dir)
if (!dir.exists(work_dir)) dir.create(work_dir)


# Getting JavaStics
# h <- curl::new_handle()
# curl::handle_setopt(h, username = "plecharpent")
# curl::handle_setopt(h, password = "vi8,gife")
# file_url <- "https://w3.avignon.inra.fr/forge/attachments/download/2019/JavaSTICS-1.41-stics-9.1.zip"
# curl::curl_download(url = file_url, handle = h, destfile = file.path(work_dir,"JavaSTICS-1.41-stics-9.1.zip"), mode = "wb")

javastics <- file.path(work_dir, params$javastics)

if (!dir.exists(javastics)) dir.create(javastics)

if (!length(grep("JavaStics.exe",list.files(javastics)))) {
  system(paste("curl -u plecharpent:vi8,gife https://w3.avignon.inra.fr/forge/attachments/download/2019/JavaSTICS-1.41-stics-9.1.zip --output", file.path(work_dir,"JavaSTICS-1.41-stics-9.1.zip")))
  
  unzip(file.path(work_dir,"JavaSTICS-1.41-stics-9.1.zip"), exdir = javastics)
}

workspace_xml <- file.path(work_dir, params$workspace_xml)
workspace_txt <- file.path(work_dir, params$workspace_txt)
ex_lines <- params$lines
```


```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.lines = 5)
#source("static_help.R")
```




## SticsRFiles package

### Description
The goal of SticsRFiles is to perform manipulations of the Stics model files either on `XML files` (used by the JavaStics GUI) or on `text files` used by the model fortran executable.

### Content of the demo
We are focusing in this demo only on `XML files` manipulations and their transformation to `text files`.

Some of the main functions are presented here with possible interactive run directly inside the document.

### Package loading and content

```{r, echo = TRUE}
library(SticsOnR)
library(SticsRFiles)
```

```{r, echo=FALSE, warning=FALSE, results='hide'}
# Generating html help files, one for each function
if(params$gen_help) {
  help_dir <- file.path(work_dir, "SticsRFiles_help")
  if (!dir.exists(help_dir)) dir.create(help_dir)
  SticsRFiles:::static_help(pkg = "SticsRFiles", 
                            overwrite = FALSE, 
                            out_dir = help_dir)
}
```

### Tutorial directory
Either all files downloaded files or generated ones while executing this tutorial will be stored under the user home directory (i.e. /home/username for linux or C:\\Users\\username\\Documents).

For the current demo session the full path is: **`r work_dir`**.

## Function get_usms_list

### Aim
> _**Getting a list of `usms` names from a usms.xml file of a workspace**_

<!-- FOR DISPLAYING a part of the online help
FROM Usage to Value -->

```{r, results="asis"}
lines <- SticsRFiles:::get_from_help(file.path(help_dir, "get_usms_list.html"))
cat( paste( lines , collapse="\n" ) )
```

### Example
```{r, echo = TRUE}
# Path used for workspace
workspace_xml
# Getting usms list
get_usms_list(usm_path = file.path(workspace_xml, "usms.xml"))
```
#### Getting usms list using a filter name
```{r get_usms_list_1, exercise=TRUE}
get_usms_list(usm_path = file.path(workspace_xml, "usms.xml"), name = "gar")

```

#### Getting usms list using filter names
```{r get_usms_list_2, exercise=TRUE}
get_usms_list(usm_path = file.path(workspace_xml, "usms.xml"), name = c("Sugar", "pe"))

```


## Function `get_param_info`

### Aim
> _**Find on the overall model parameters information a kwown parameter, with a possible search using a partial match string.**_

<!-- FOR DISPLAYING a part of the online help
FROM Usage to Value -->

```{r, results="asis"}
lines <- SticsRFiles:::get_from_help(file.path(help_dir, "get_param_info.html"))
cat( paste( lines , collapse="\n" ) )
```

### Example
```{r, echo = TRUE}
get_param_info("albedo", version = "V9.1")
```


### Use cases examples

#### Finding parameters names containing a keyword

```{r find_param_names_1, exercise=TRUE}
get_param_info(parameter = "hum", version = "V9.1")

get_param_info(parameter = "prof", version = "V9.1")
```

#### Finding parameters names containing several keywords

```{r find_param_names_2, exercise=TRUE}
get_param_info(parameter = c("profsem", "profhum"), version = "V9.1")

```
<!--
#### The function `exist_param_xml` may also be used for checking if parameters names exist

```{r exist_param_xml, exercise = TRUE, warning = FALSE}
exist_param_xml(name = c("profsem", "profhum"), version = "V9.1")

```
-->

## Function get_param_xml

```{r sols_path, echo=FALSE}
xml_path <- system.file("extdata/xml/examples/V9.1", package = "SticsRFiles")
xml_soil <- file.path(workspace_xml,"sols.xml")
```

### Aim
> _**Find parameters information using their name or a part of it.
> For a file or a set of files**_

<!-- FOR DISPLAYING a part of the online help
FROM Usage to Value -->


```{r, results="asis"}
lines <- SticsRFiles:::get_from_help(file.path(help_dir, "get_param_xml.html"))
cat( paste( lines , collapse="\n" ) )
```

### Example

Getting parameter values from a soils parameter file
```{r, echo = TRUE}
# File containing soils parameters
xml_soil

get_param_xml(xml_file = xml_soil, param_name = "albedo")
```


### Use cases examples

#### Getting parameters values for the parameter "argi" for all soils and the "solble" soil


```{r get_param_xml_1, exercise=TRUE}
get_param_xml(xml_file = xml_soil, param_name = "argi")

get_param_xml(xml_file = xml_soil, param_name = "argi", select = "sol", value = "solble")
```
#### Getting parameters values for the parameters "argi" and "norg" for all soils and the "solble" soil

```{r get_param_xml_2, exercise=TRUE}

get_param_xml(xml_file = xml_soil, param_name =  c("argi", "norg"))

#get_param_xml(xml_files = xml_soil, param_names =  c("argi", "norg"), select = "sol", value = "solble")
```

#### Getting soil parameters names and playing with `get_param_xml` using parameters names

```{r get_param_xml_3, exercise=TRUE, warning = FALSE}

get_param_names_xml(xml_file = xml_soil)

get_param_xml(xml_file = xml_soil, param_name = "paramtest", select = "sol", value = "solble")
```



## Function set_param_xml


### Aim
> _**Setting parameter value or a vector of**_


```{r, results="asis"}
lines <- SticsRFiles:::get_from_help(file.path(help_dir, "set_param_xml.html"))
cat( paste( lines , collapse="\n" ) )
```

### Example
```{r, echo = FALSE, results='hide'}
xml_gen_src <- file.path(xml_path,"param_gen.xml")
xml_gen <- file.path(workspace_xml,"param_gen.xml")
file.copy(from = xml_gen_src, xml_gen, overwrite = TRUE)
```

```{r, echo = TRUE}
# File containing general parameters
xml_gen

set_param_xml(xml_gen, param_name = "tnitopt", param_value = 29.5, overwrite = TRUE )
# Controlling the written values
get_param_xml(xml_gen, param_name = "tnitopt" )

```


### Use cases examples
```{r}
set_param_xml(xml_soil, param_name = "argi", param_value = 23.1, select = "sol", value = "solble", overwrite = T )
xml_tec <- file.path(workspace_xml,"Mais_tec.xml")
jul <- c(178, 185, 193, 198, 200, 204, 207, 211, 214, 218, 221, 232, 239, 249, 257, 264)
val <- c(20, 24, 29, 29, 28, 31, 21, 21, 23, 18, 22, 16, 16, 30, 29, 20)
set_param_xml(xml_tec, param_name = c("julapI_or_sum_upvt", "amount"), param_value = list(jul, val), overwrite = T )
```

#### Setting a simple parameter value
```{r set_param_xml_1, exercise=TRUE, warning = FALSE, exercise.lines = 10}
param <- "codeactimulch"
# Controlling the initial value
get_param_xml(xml_gen, param_name = param )

set_param_xml(xml_gen, param_name = param, param_value = 2, overwrite = TRUE )
# Controlling the written value
get_param_xml(xml_gen, param_name = param )
```

#### Setting a simple parameter value with selection
```{r set_param_xml_2, exercise=TRUE, warning = FALSE, exercise.lines = 10}
param <- "argi"
# Controlling initial value
get_param_xml(xml_soil, param_name = param, select = "sol", value = "solble")

# Using a conditional selection
set_param_xml(xml_soil, param_name = param, param_value = 33, select = "sol",
              value = "solble", overwrite = T )

# Controlling written values
get_param_xml(xml_soil, param_name = param, select = "sol", value = "solble")
```

#### Setting multiple values

```{r set_param_xml_3, exercise=TRUE, warning = FALSE, exercise.lines = 10}

# Controlling initial value
get_param_xml(xml_tec, param_name = c("julapI_or_sum_upvt", "amount"))

set_param_xml(xml_tec, param_name = c("julapI_or_sum_upvt", "amount"), param_value = list(200:215,20:35), overwrite = T )

# Controlling written values
get_param_xml(xml_tec, param_name = c("julapI_or_sum_upvt", "amount"))
```


## Function gen_usms_xml2txt


### Aim
> _**Generating one or several usms directories from a javastics workspace**_


```{r, results="asis"}
lines <- SticsRFiles:::get_from_help(file.path(help_dir, "gen_usms_xml2txt.html"))
cat( paste( lines , collapse="\n" ) )
```

### Example

```{r, echo = FALSE}
# Generating once, if all dirs exist
l <- file.path(workspace_xml, get_usms_list(usm_path = file.path(workspace_xml, "usms.xml"))[[1]])
all_exists <- all(file.exists(l))

if (!all_exists) {
  value <- gen_usms_xml2txt(javastics_path = javastics, workspace_path = workspace_xml)
}
```


```{r, echo = TRUE, eval = FALSE}
# Workspace used as usms XML files source
workspace_xml

# Generating all the usms input directories
value <- gen_usms_xml2txt(javastics_path = javastics, workspace_path = workspace_xml)

value
## $usms_paths
##  [1] "SugarCane"            "potato"               "banana"              
##  [4] "sorghum"              "barley"               "sugarbeet"           
##  [7] "wheat"                "maize"                "soybean"             
## [10] "lettuce"              "tomato"               "DurumWheat"          
## [13] "rapeseed"             "sunflower"            "grass"               
## [16] "BareSoil"             "demo_Wheat1"          "demo_BareSoil2"      
## [19] "demo_maize3"          "DurumWheat_4years"    "maize_4years"        
## [22] "strawberry"           "pea"                  "vine"                
## [25] "fescue"               "flax"                 "intercrop_pea_barley"
## [28] "timothy"              "DurumWheat_snow"      "Turmeric"            
## [31] "cc_BristleOat"        "cc_mustard"           "cc_ItalianRyegrass"  
## [34] "cc_vetch"             "cc_CrimsonClover"     "proto_rice"          
## 
## $files_path
##  [1] "/tmp/Demo-SticsRPacks/JavaSTICS-1.41-stics-9.1/example/climat.txt"     
##  [2] "/tmp/Demo-SticsRPacks/JavaSTICS-1.41-stics-9.1/example/param.sol"      
##  [3] "/tmp/Demo-SticsRPacks/JavaSTICS-1.41-stics-9.1/example/ficini.txt"     
##  [4] "/tmp/Demo-SticsRPacks/JavaSTICS-1.41-stics-9.1/example/ficplt1.txt"    
##  [5] "/tmp/Demo-SticsRPacks/JavaSTICS-1.41-stics-9.1/example/fictec1.txt"    
##  [6] "/tmp/Demo-SticsRPacks/JavaSTICS-1.41-stics-9.1/example/station.txt"    
##  [7] "/tmp/Demo-SticsRPacks/JavaSTICS-1.41-stics-9.1/example/new_travail.usm"
##  [8] "/tmp/Demo-SticsRPacks/JavaSTICS-1.41-stics-9.1/example/tempopar.sti"   
##  [9] "/tmp/Demo-SticsRPacks/JavaSTICS-1.41-stics-9.1/example/tempoparv6.sti" 
## [10] "/tmp/Demo-SticsRPacks/JavaSTICS-1.41-stics-9.1/example/ficplt2.txt"    
## [11] "/tmp/Demo-SticsRPacks/JavaSTICS-1.41-stics-9.1/example/fictec2.txt"    
## 
## $copy_status
##  [1] TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE
## [16] TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE TRUE
## [31] TRUE TRUE TRUE TRUE TRUE TRUE
## 
## $obs_copy_status
##  [1]  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE
## [13]  TRUE  TRUE  TRUE  TRUE FALSE FALSE FALSE FALSE FALSE  TRUE  TRUE  TRUE
## [25]  TRUE  TRUE FALSE  TRUE FALSE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE


```


### Use cases examples
```{r, echo = FALSE}
#value <- gen_usms_xml2txt(javastics_path = javastics, workspace_path = workspace_xml, target_path = workspace_txt)
```

#### Generating directories for a subset of usms
```{r , warning = FALSE, echo = TRUE}
# Workspace used as usms XML files source
workspace_xml

# Existing usms in the xml workspace directory
get_usms_list(usm_path = file.path(workspace_xml, "usms.xml"))
```

```{r gen_usms_1, exercise=TRUE, warning = FALSE}
value <- gen_usms_xml2txt(javastics_path = javastics, workspace_path = workspace_xml, usms_list = c("wheat", "pea"))

value

```

#### Generating files in a specific directory
```{r gen_usms_2, exercise=TRUE, warning = FALSE, exercise.lines = 10}
# Directory where to generate files
workspace_txt

value <- gen_usms_xml2txt(javastics_path = javastics, 
                          workspace_path = workspace_xml,
                          target_path = workspace_txt, 
                          usms_list = c("wheat", "pea" ))
value

```


